\documentclass[11pt]{article}

\usepackage{graphicx}

\usepackage{amsmath,amsfonts,amssymb}

\usepackage{color}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=black,filecolor=blue,urlcolor=blue]{hyperref}
\newcommand{\todo}[1]{{\bf \color{blue} [#1]}}
\newcommand{\alert}[1]{{\bf \color{red} [#1]}}
\newcommand{\newsection}{\clearpage\newpage}

\setlength{\textwidth}{6.2in}
\setlength{\oddsidemargin}{0.3in}
\setlength{\evensidemargin}{0in}
\setlength{\textheight}{8.7in}
\setlength{\voffset}{-.7in}
\setlength{\headsep}{26pt}

\input{RJLmacros}

\graphicspath{{./}}


% indicate which bp*.tex files to include: (comment out to include all)
\includeonly{bp7}


\begin{document}

\begin{center}
{{\Large\bf
GeoClaw Results for \\
NTHMP Tsunami Benchmark Problems}
\vskip 5pt
Version of \today}
\end{center}

% NTHMP Model validation workshop : draft reporting template

\vskip 15pt
\noindent
{\bf Team:}  GeoClaw Tsunami Modeling Group, University of Washington

\vskip 5pt
\noindent
{\bf Geographic area:}  The west coast, with a focus on Washington State,
including Puget Sound, the San Juan Island group, and other islands in the
Strait of Juan de Fuca.

\vskip 5pt
\section{Introduction}

(Describe your NTHMP funded effort regarding inundation mapping, i.e., which
area this covers, which types of tsunami sources are affecting your area and
hence which benchmark problems in the list you will be addressing and why.)

Funding is expected from NTHMP partner FEMA’s Risk Mapping, Assessment, and
Planning (Risk MAP) Program, as part of the BakerAECOM coastal engineering
effort for the FEMA California Coastal Analysis and Mapping (CCAMP) Project.
Funding will also be sought to provide Washington State with tsunami
modeling and mapping in support of tsunami hazard assessment and emergency
management planning and education.  Tsunami sources in these geographic
areas include earthquakes and landslides, and we will therefore address
benchmark problems that deal with these sources.

\section{Model description}
\todo{
(Describe and give relevant references to the model(s) you are using for
your tsunami source, propagation, and coastal impact/inundation simulations.
As a minimum, provide general information on the type of models you are
using, including related restriction/assumptions. Refer to (preferably peer
reviewed) papers where your model equations and other relevant validation
and/or case studies for tsunami or otherwise have been published.)
}

The GeoClaw software is described in some detail in
\cite{BergerGeorgeLeVequeMandli:awr10,LeVequeGeorgeBerger:an11} and in the
on-line documentation \cite{geoclaw-doc}.  GeoClaw is a variant of the
Clawpack open source
software \cite{claw.org-doc} that LeVeque and collaborators have
been developing since 1994.  The ``wave-propagation algorithms'' used in
this software are described in great detail in the textbook
\cite{rjl:fvmhp} and in several journal publications
\cite{rjl:advect,rjl:waveprop,jol-rjl:3d}.

Adaptive mesh refinement (AMR) has been
incorporated since the inception of this software, 
in joint work with Marsha Berger at the
Courant Institute of Mathematical Sciences, one of the foremost authorities
on block structured AMR technology, often referred to as
Berger-Colella-Oliger style AMR \cite{berger-colella, berger-oliger}.  The
AMR algorithms in Clawpack are described in detail in
\cite{mjb-rjl:amrclaw}.  

GeoClaw was initiated in the PhD work of David George
\cite{dgeorge:phd,rjl-george:catalina04a,dg-rjl:tsunami06,George2008} and
was originally called TsunamiClaw.  It has more recently been extended to
handle other geophysical flows such as storm surge \cite{mandli:?}
dam break problems \cite{George:Malpasset}, and debris flows
\cite{GeorgeIverson2011}.

The GeoClaw model solves the two-dimensional nonlinear shallow water equations 
using high-resolution finite volume methods.  Values of $h$, $hu$, and $hv$
in each grid cell represent cell averages of the depth and momentum
components.  With flat bathymetry, the methods are exactly conservative for
both mass and momentum, and conserve mass for arbitrary bathymetry when used
on a fixed grid. They do not exactly conserve mass during regridding when
adaptive mesh refinement is used (discussed further below).

These methods are based on Godunov's method, which means that at each cell
interface a one-dimensional Riemann problem is solved normal to the edge,
which reduces to a one-dimensional shallow water model with piecewise
constant initial data,
with left and right values given by the cell averages on each side.
The jump in bathymetry between the cells is incorporated into the Riemann
solution in a manner that makes the method ``well balanced'': the steady
state of the ocean at rest is exactly maintained.  This is done using the
``f-wave formulation'' of the wave propagation method, as discussed in
\cite{BaleLeVequeEtAl2002,George2008,rjl:fvmhp,rjl:wbfwave10}.


Godunov's method consists of solving the Riemann problem and using the
resulting wave structure to update cell averages in the adjacent finite
volume cells.  In practice an approximate Riemann solution is used, which
reduces to the standard Roe solver \cite{rjl:fvmhp,roe:rs}
in f-wave form in general, but is modified
to also handle dry states in order to model inundation \cite{George2008}.

\subsection{Other validation studies}

Validation studies of the GeoClaw software are presented in the following
papers:

\begin{itemize}
\item In \cite{BergerGeorgeLeVequeMandli:awr10} a test problem is used that
consists of a radially symmetric ocean with a continental shelf, and
a radially symmetric initial
hump of water at the center.  At one position along the coast a 
island is placed on the shelf.  Several gauges are located near the island.
If the island is rotated to a different position on the coast the results at
the gauges should be identical.  Numerically they will not be identical due
to differences in how the coast intersects the Cartesian grid.  Very similar
results are obtained for locations near an axis and near the diagonal, both
in gauge results and in plots of the surface and inundation.  This is true
even when AMR is used to concentrate fine grids only in the direction
towards the island.  In this paper the Chile 2010 tsunami is also used as a
test problem and good agreement with measured results at DART Buoy 21413??
are obtained, on two different grid resolutions to test convergence.

\item Further tests of this same problems are presented in
\cite{LeVequeGeorgeBerger:an11}.  

\item In \cite{mjb-dac-ch-rjl:amrsphere09} tests are performed for a
tsunami-like wave propagating on the full sphere, using a novel mapped grid
that covers the sphere with a logically rectangular finite volume grid.
Tests are presented in which the bathymetry and the initial conditions are
axisymmetric so that the solution should remain so and can be compared to
fine grid one-dimensional simulations.  
\end{itemize} 




\section{Benchmark results}
\todo{(For each selected (PMEL) benchmark problem, please document how you defined
your model input data, grid and various other parameters to simulate each
problem. 

Provide results as figures (and raw data) following future formatting
instructions, to the relevant site and/or coordinator. Also summarize your
results as simple error indicators of the performance of the model (e.g.,
mean, max, rms errors).)}

In the sections below,
PMEL-135 refers to \cite{SynolakisBernard:pmel135}.

\include{bp1}
\include{bp2}
\include{bp3}
\include{bp4}
\include{bp5}
\include{bp6}
\include{bp7}
\include{bp8}
\include{bp9}
\include{bp10}
\include{bp11}


\bibliographystyle{plain}
\bibliography{references}
\end{document}

